Study of the hadronization of charged pions

### 2019-03-22 chap5.tex
minor edits

The acceptance correction consists of applying weights to the experimental data
event by event to correct for any detection inefficiencies of used detectors.
Incidentally, it also corrects for other small detection issues, such as 
misidentification and re-scattering on detector material. The quality 
of the correction depends on the ability of the simulation to reproduce the 
experiment, the accumulated statistics, and the size of interfering effects, such 
as the bin migration. For this correction, we applied the method from the 
approved and published analysis from the same eg2 data~\cite{ElFassi:2008}.

\paragraph{Simulation} ~\\
To correct for acceptance effects, we simulated a total of 100 million events 
per target ($^2$H, C, Fe and Pb) using the PYTHIA \cite{Sjostrand:2006za} 
event generator that was slightly modified to include Fermi motion effects. The 
generated events are processed by CLAS softwares (GSIM, GPP and user\_ana) 
to simulate the detector and the reconstruction process in a fashion similar 
to that done with the experimental data.
% TODO GPP factors

The simulated data are processed like the experimental data by 
applying the cuts described in the section \ref{sec:pid}. Overall the 
simulation reproduces quite well the detector response, yet two issues might 
affect us and have to be understood. First, the efficiency of the CC is 
overestimated in the simulation. On the electron side the signal is a little stronger in the simulation (11 
photo-electrons) compared to experimental data (8 photo-electrons), but this 
feature should not affect us too much, because we are cutting only the tail of the 
distribution in both cases. Second, since in the simulation the beam is perfectly aligned with CLAS, the vertex cuts do not need to be shifted from one sector to an other. Therefore, we do not apply the shift from table \ref{tab:vertex} to the simulated data; the resulting cuts are shown in figure~\ref{simvertex}.

@@

The kinematical distributions from the simulation are compared to the 
experimental ones. This is important for the acceptance correction, to see 
whether or not, we can integrate over any variables in this correction.
Comparisons between simulated and experimental data are shown in figures 
\ref{fig:compNuQ2} to \ref{fig:compPhih}. The agreement is reasonable, but not 
perfect. The differences are due to the PYTHIA simulation, which is not including some 
physical effects, such as radiative and diffractive processes.

@@

The weights ($\omega_{\pi^\pm}(\nu,x_{Bj},p_h,t)$) are extracted from 
the simulation using equation \ref{eq:acc}. However, these coefficients should not be applied to data before resolving the issues with bins that have either
\begin{itemize}
 \item too large weight values, i.e., very small acceptance, or
 \item very few reconstructed events and very large statistical errors, or
 \item a large fraction of events originally generated in another bin (bin migration issue).
\end{itemize}
To eliminate these bins we apply the following cuts:
\begin{equation} \label{eq:AccL1}
\left ({\delta \omega \over \omega}\right ) ^2 \times R_c \times \omega < 2,
\end{equation}
and
\begin{equation} \label{eq:AccL2}
N_{rec} > 4,
\end{equation}
where $\delta \omega / \omega$ is the relative error on the weight, and $R_c$ is the 
proportion of reconstructed events initially generated in an other kinematical bin. 
Figure~\ref{fig:AccCoef} shows $\omega$ distributions after applying these cuts. Around 
900 bins remain for both targets, but we noticed that these weights are much larger for 
$\pi^-$ compared to $\pi^+$, as an indication of a larger correction for the inbent 
particles. The motivation for using such a cut based on the product of different unrelated
values is actually directly linked to this problem. The $\pi^-$ acceptance drops very 
strongly at low \pt leading to very high weights, however there is no problem of bin migration
or low statistics because the generated yield is very high in this region. We found that 
the best way
to keep these bins while discarding bins where the weight might be slightly smaller but which 
have issues with bin migration and statistics was to apply this set of cuts.

In order to correct for these droped bins and reduce the sensitivity to the arbitrary in
the choice of the weights' cuts, we corrected the data with an additional two dimensional 
reweighting factor as:
\begin{equation}
\text{Acc}_2(\nu,p_h) = {\sum_{rec} \omega (\nu,x_{Bj},p_h,t)\over N_{gen}(\nu,p_h)}.
\end{equation}
where $\omega$ is the previously calculated weight or a 1 for a bin that was excluded by the 
cuts of equation \ref{eq:AccL1} and \ref{eq:AccL2}. One of the criteria for setting the limits 
in these equation is in fact to keep this reweighting factors, $\omega_2 = 1/\text{Acc}_2$, at 
the level of a few percents. This indicates that the previous method only missed a small 
fraction of the event in the first place and is a good indication of the sanity of the method.

@@

Figure \ref{fig:AcceptPlots}, where only statistical errors are 
represented, shows the new results after applying the acceptance correction.
The acceptance correction effects appear to be in the order of $10$\% for 
multiplicity ratios and $30$\% for the transverse momentum broadening. This is a surprisingly significant correction with regard to a few centimeters separating the two targets. The reason for such a large correction is that the small shift in the $z$-vertex position changes slightly the minimum angle of the detector in the region where most hadron production is concentrated, {\it i.e.} at low \pt close to the beam line. The $\pi^-$s were more affected due to the inbending magnetic field. Indeed, the low \pt $\pi^-$s had very low acceptance which led to large correction factors. The large spread in weights makes the errors on \dpt very large, and the related results difficult to exploit. However, our transverse momentum broadening 
results of both pions matched each other after this correction, which is similar 
to the HERMES results (see figure~\ref{fig:her3}).

@@

The large electric charge of our nuclear targets create an electric field that 
should be taken into account at our energies. Besides the fact that the nuclear 
electric field accelerates or decelerates the incoming and outgoing charged 
particles, it has some non trivial effects such as the focusing of low energy 
particles. Aste {\it et al.}~\cite{Aste:2005wc} showed that these Coulomb 
effects could be corrected for by using an effective momentum approximation at 
large momenta ({\it i.e.}, $Q^2>.5$ GeV$^2$). As our kinematics coverage is 
beyond this limit, we decided to use this approximation on the Coulomb 
correction of our data-sets. According to Aste {\it et al.}, the best effective 
potential approximation is given by $\bar V= 0.8 V_0$, where $V_0$ is the 
electrostatic potential at the center of nuclei. The latter is evaluated as 
$V_0= -3 \alpha_{EM} Z / (2 R)$, where $\alpha_{EM}$ is the fine-structure 
constant, $Z$ is the proton number, and $R$ is the radius of the nuclei. Table 
\ref{tab:Coulomb} contains the used values of $V_0$ and $\bar V$ on different 
nuclear target's analysis. The Coulomb correction is applied to both electrons 
and charged pions before calculating the plotted kinematical variables.

@@

TODO Add figures of the effect and comment from review answers!

@@

To apply the radiative correction, we apply a weight to each events in a fashion
similar to the acceptance correction. The HAPRAD 2 code provides 
$\delta = \sigma_{obs} / \sigma_B$ at given kinematics. Table \ref{tab:RCBins} 
shows the binning used to calculate the radiative correction weights.
We chose to make a finer 
binning for the radiative corrections than for acceptance as the computing time 
to calculate weights is more reasonnable (around 1000 days of cpu time).

@@

We show in figure \ref{fig:RCexample} a selected sample of radiative correction factors
($= \delta_{Fe} / \delta_D$) in order to illustrate the expected effect on observables.
In the large majority of the cases the factors are within 5\% of 1, so we do not expect to
see a strong effect on the final results where these will be integrated. This is confirmed 
in figure \ref{} ... In the case of the transverse momentum braodening, the variations
appear to ...

@@

\begin{figure}[tbp]
\centering
\caption {Preliminary results ..., 
for multiplicity ratios (top) and transverse momentum broadening 
(bottom). Only statistical errors are shown.}
\label{fig:RCPlots}
\end{figure}

@@

>>>>

The acceptance correction consists of applying weights to the experimental data
event by event to correct for the detection inefficiencies of the detectors.
Incidentally, it also corrects for other small detection issues, such as 
misidentification and scatterings on detector material. The quality 
of the correction depends on the ability of the simulation to reproduce the 
experiment, the accumulated statistics, and the size of interfering effects, such 
as bin migration. For this correction, we applied the method from the 
approved analysis from the same eg2 data~\cite{ElFassi:2008}.

\paragraph{Simulation} ~\\
To correct for acceptance effects, we simulated a total of 100 million events 
per target ($^2$H, C, Fe and Pb) using the PYTHIA \cite{Sjostrand:2006za} 
event generator that was slightly modified to include Fermi motion effects. The 
generated events are processed by CLAS softwares (GSIM, GPP and user\_ana) 
to simulate the detector and the reconstruction process in a fashion similar 
to that done with the experimental data.
% TODO GPP factors: -a1.7 -b1.9 -c2. -f1. -R41147 (likely the one used)

The simulated data are processed like the experimental data by 
applying the cuts described in the section \ref{sec:pid}. Overall the 
simulation reproduces quite well the detector response, yet two issues might 
affect us and have to be understood. First, the efficiency of the CC is 
overestimated in the simulation. On the electron side the signal is a little stronger in the simulation (11 
photo-electrons) compared to experimental data (8 photo-electrons), but this 
feature should not affect us too much, because we are cutting only the tail of the 
distribution in both cases. Second, since in the simulation the beam is perfectly 
aligned with CLAS, the vertex cuts do not need to be shifted from one sector to 
another. Therefore, we do not apply the shift from table \ref{tab:vertex} to the 
simulated data; the resulting cuts are shown in figure~\ref{simvertex}.

@@

The kinematical distributions from the simulation are compared to the 
experimental ones. This is important for the acceptance correction, to see 
whether or not, we can integrate over any variables in this correction.
Comparisons between simulated and experimental data are shown in figures 
\ref{fig:compNuQ2} to \ref{fig:compPhih}. The agreement is reasonable, but not 
perfect. The differences are due to the PYTHIA simulation, which is not including some 
physical effects, such as radiative effects and diffractive processes.

@@

The weights ($\omega_{\pi^\pm}(\nu,x_{Bj},p_h,t)$) are extracted from 
the simulation using equation \ref{eq:acc}. However, these coefficients should not be applied to data before resolving the issues with bins that have either
\begin{itemize}
 \item too large weight values, i.e., very small acceptance, or
 \item very few reconstructed events and very large statistical errors, or
 \item a large fraction of events originally generated in another bin (bin migration issue).
\end{itemize}
To eliminate these bins we apply the following cuts:
\begin{equation} \label{eq:AccL1}
\left ({\delta \omega \over \omega}\right ) ^2 \times R_c \times \omega < 2,
\end{equation}
and
\begin{equation} \label{eq:AccL2}
N_{rec} > 4,
\end{equation}
where $\delta \omega / \omega$ is the relative error on the weight, and $R_c$ is the 
proportion of reconstructed events initially generated in another kinematical bin. 
Figure~\ref{fig:AccCoef} shows $\omega$ distributions after applying these cuts. Around 
900 bins remain for both targets, but we noticed that these weights are much larger for 
$\pi^-$ compared to $\pi^+$, as an indication of a larger correction for the inbent 
particles. The motivation for using such a cut based on the product of different unrelated
values is actually directly linked to this problem. The $\pi^-$ acceptance drops very 
strongly at low \pt leading to very high weights, however there is no problem of bin migration
or low statistics because the generated yield is very high in this region. We found that 
the best way
to keep these bins while discarding bins where the weight might be slightly smaller but which 
have issues with bin migration and statistics was to apply the cuts presented above.

This process can create a bias, if for instance low $Q^2$ bins are more often dropped.
So, in order to correct for these bins and reduce the sensitivity to the arbitrary in
the choice of the cuts, we corrected the data with an second two dimensional 
acceptance correction as:
\begin{equation}
\text{Acc}_2(\nu,p_h) = {\sum_{rec} \omega (\nu,x_{Bj},p_h,t)\over N_{gen}(\nu,p_h)}.
\end{equation}
where $\omega$ is the previously calculated weight or a 1 for a bin that was excluded by the 
cuts of equation \ref{eq:AccL1} and \ref{eq:AccL2}. One of the criteria for setting the limits 
in these equation is in fact to keep this reweighting factors, $\omega_2 = 1/\text{Acc}_2$, at 
the level of a few percents. This indicates that the previous method only missed a small 
fraction of the event in the first place and is a good indication of the sanity of the method.

@@

Figure \ref{fig:AcceptPlots}, where only statistical errors are 
represented, shows the new results after applying the acceptance correction.
The acceptance correction effects appear to be of the order of $10$\% for 
multiplicity ratios and $30$\% for the transverse momentum broadening. This 
is a surprisingly significant correction with regard to a few centimeters 
separating the two targets. The reason for such a large correction is that 
the small shift in the $z$-vertex position changes slightly the minimum 
angle of the detector in the region where most hadron production is 
concentrated, {\it i.e.} at low \pt close to the beam line. The 
$\pi^-$s were more affected due to the inbending magnetic field. Indeed, 
the low \pt $\pi^-$s had very low acceptance which led to large correction 
factors. The large spread in weights makes the errors on \dpt rather large. 
However, the transverse momentum broadening 
results of both pions matched each other better after this correction, which is similar 
to the HERMES results (see figure~\ref{fig:her3}).

@@

The large electric charge of our nuclear targets create an electric field that 
should be taken into account at our energies. Besides the fact that the nuclear 
electric field accelerates or decelerates the incoming and outgoing charged 
particles, it has some non trivial effects such as the focusing of low energy 
particles. Aste {\it et al.}~\cite{Aste:2005wc} showed that these Coulomb 
effects could be corrected for by using an effective momentum approximation at 
large momenta ({\it i.e.}, $Q^2>.5$ GeV$^2$). As our kinematics coverage is 
beyond this limit, we decided to use this approximation on the Coulomb 
correction of our data-sets. According to Aste {\it et al.}, the best effective 
potential approximation is given by $\bar V= 0.8 V_0$, where $V_0$ is the 
electrostatic potential at the center of nuclei. The latter is evaluated as 
$V_0= -3 \alpha_{EM} Z / (2 R)$, where $\alpha_{EM}$ is the fine-structure 
constant, $Z$ is the proton number, and $R$ is the radius of the nuclei. Table 
\ref{tab:Coulomb} contains the used values of $V_0$ and $\bar V$ on different 
nuclear target's analysis.

@@

The Coulomb correction is applied to incoming and outgoing electrons as well as
produced charged pions. The effect of the correction is shown in figure \ref{fig:CCPlots},
we observe that it is small in most of the phase space, but by slightly shifting 
the $z$ distribution it affects the normalization at a couple of percent level. A
point to point effect is also visible for low $z$ pions. Note that in general, there
is also a focussing factor in Coulomb corrections, but they happen to cancel out in
our specific observables.

\begin{figure}[tbp]
\centering
\includegraphics[width=7.4cm] {chap5-fig/b_RvA.png} 
\includegraphics[width=7.4cm] {chap5-fig/b_RvZ.png} 
\includegraphics[width=7.4cm] {chap5-fig/b_PvA.png} 
\includegraphics[width=7.4cm] {chap5-fig/b_PvQ2.png} 
\caption {Preliminary results are compared with the acceptance corrected ones, 
for multiplicity ratios (top) and transverse momentum broadening 
(bottom). Only statistical errors are shown.}
\label{fig:CCPlots}
\end{figure}

@@

To apply the radiative correction, we apply a weight to each events in a fashion
similar to the acceptance correction. The HAPRAD 2 code provides 
$\delta = \sigma_{obs} / \sigma_B$ at given kinematics. Table \ref{tab:RCBins} 
shows the binning used to calculate the radiative correction weights.
We chose to make a finer 
binning for the radiative corrections than for acceptance as the computing time 
to calculate weights is more reasonnable ("only" around 1000 days of CPU time).

@@

We show in figure \ref{fig:RCexample} a selected sample of radiative correction factors
($= \delta_{Fe} / \delta_D$) in order to illustrate the expected effect on observables.
In the large majority of the cases the factors are within 5\% of 1. Therefore, we do not expect to
see a strong effect on the final results where these will be integrated. This is confirmed 
in figure \ref{fig:RCPlots}, where the radiative correction appears limited to a few percents. 
In the case of the transverse momentum braodening, the variations appear to be similar in size 
and not too worrisome.

@@

\begin{figure}[tbp]
\centering
\includegraphics[width=7.4cm] {chap5-fig/b_RvA.png} 
\includegraphics[width=7.4cm] {chap5-fig/b_RvZ.png} 
\includegraphics[width=7.4cm] {chap5-fig/b_PvA.png} 
\includegraphics[width=7.4cm] {chap5-fig/b_PvQ2.png} 
\caption {Preliminary results including radiative corrections
compared to the ones without, 
for multiplicity ratios (top) and transverse momentum broadening 
(bottom). Only statistical errors are shown.}
\label{fig:RCPlots}
\end{figure}

@@

\subsubsection{Pion ratio from HERMES}
